---
title: "mIHC Report"
author: "Wes Horton"
date: "`r Sys.Date()`"
output: 
  html_document:
      code_folding: hide
      toc: true
      toc_depth: 3
      toc_float: TRUE
      number_sections: TRUE
      css: doc.css
params:
  reportID: "empty"
  proj: "empty"
  cellFile: "empty"
  funcFile: "empty"
  metaFile: "empty"
  colorFile: "empty"
  configFile: "empty"
  reportCol: "empty"
  idCol: "empty"
  sampleCol: "empty"
---

```{r loadFxn, echo = F}
### For some reason, loading this from package isn't working right now.
tempMergeDTs <- function(data_lsdt, mergeCol_v, keepCol_v = NULL, ...) {
  #' Merge many data.tables together
  #' @description Take many data.tables and merge on and ID column, extracting a single column from each data.table as the column of interest
  #' @param data_lsdt list of data.tables to merge
  #' @param mergeCol_v which column from all of the data.tables to use to merge
  #' @param keepCol_v which column from all of the data.tables to use as the column of interest. If NULL, use all columns
  #' @param ... extra parameters passed to merge
  #' @return data.table with ncol == length(data_lsdt) + 1. Column names are names of list, or defaults to V1, V2,...
  #' @export
  
  ## Grab extra arguments
  extraParams_lsv <- list(...)
  
  ## Handle extra arguments
  if (!is.null(extraParams_lsv$all)){
    all_v <- extraParams_lsv$all
  } else {
    all_v <- T
  } # fi
  
  if (!is.null(extraParams_lsv$sort)){
    sort_v <- extraParams_lsv$sort
  } else {
    sort_v <- F
  } # fi
  
  ## If keepCol_v is NULL, grab all other columns
    if (is.null(keepCol_v)){
    keepCol_v <- colnames(data_lsdt[[1]])[-which(colnames(data_lsdt[[1]]) %in% mergeCol_v)]
  } # fi
  
  ## Create initial table by extracting the 2 columns of interest from the rest
  merge_dt <- data_lsdt[[1]][,mget(c(mergeCol_v, keepCol_v))]

  ## Create initial column names (first check if list has names and add if not)
  if (is.null(names(data_lsdt))) {
    names_v <- paste("V", 1:length(data_lsdt))
    names(data_lsdt) <- names_v
  } # fi
  
  if (length(keepCol_v) > 1){
    colNames_v <- c(mergeCol_v, paste(names(data_lsdt)[1], keepCol_v, sep = "_"))
  } else {
    colNames_v <- c(mergeCol_v, names(data_lsdt)[1])
  } # fi
  
  for (i in 2:length(data_lsdt)) {
    
    ## This is new (2018-10-10) - need to make new keepCol_v if the data.tables don't have same columns
    if (!keepCol_v %in% colnames(data_lsdt[[i]])) {
      keepCol_v <- colnames(data_lsdt[[i]])[-which(colnames(data_lsdt[[i]]) %in% mergeCol_v)]
    } # fi
    
    ## Merge
    merge_dt <- merge(merge_dt,
                      data_lsdt[[i]][,mget(c(mergeCol_v, keepCol_v))],
                      by = mergeCol_v,
                      all = all_v, sort = sort_v)
    ## Update column names
    if (length(keepCol_v) > 1){
      colNames_v <- c(colNames_v, paste(names(data_lsdt)[i], keepCol_v, sep = "_"))
    } else {
      colNames_v <- c(colNames_v, names(data_lsdt)[i])
    } # fi
    
    ## Rename columns
    colnames(merge_dt) <- colNames_v
  } # for i
  return(merge_dt)
} # mergeDTs

```

```{r libraries, echo = F, message=F, warning=FALSE}
options(width=20)
library(readxl)
library(data.table)
library(wrh.rUtils)
library(ggplot2)
library(ggpubr)
library(gridExtra)
library(plotly)
library(tools)
library(kableExtra)
library(mIHC)

```

```{r loadData, echo = F}

### Load data
reportID_v <- params$reportID
proj_v <- params$proj

cell_dt <- readCSVOrExcel(params$cellFile)
func_lsdt <- readAllExcel(params$funcFile)
meta_dt <- readCSVOrExcel(params$metaFile)
color_dt <- readCSVOrExcel(params$colorFile)
config_dt <- readCSVOrExcel(params$configFile)

reportCol_v <- params$reportCol
idCol_v <- params$idCol
sampleCol_v <- params$sampleCol

```

```{r wrangleData, echo = F}

### Subset both data and metadata for the given report ID, if provided
### Also subset meta.data columns
if (!is.null(reportID_v)) {
  subMeta_dt <- unique(meta_dt[get(reportCol_v) == reportID_v, mget(c("reportID", "facet", "sampleName"))])
} 

### Subset data
subCell_dt <- cell_dt[get(sampleCol_v) %in% subMeta_dt[[sampleCol_v]],]
subFunc_lsdt <- func_lsdt[subMeta_dt[[sampleCol_v]]]

### Check that they have same rows
if (nrow(subMeta_dt) != nrow(subCell_dt)) stop("Unequal rows of cell data and meta. Check inputs.")
if (nrow(subMeta_dt) != length(subFunc_lsdt)) stop("Unequal rows of functional data and meta. Check inputs.")

### Check that sample column is the same
metaClass_v <- class(subMeta_dt[[sampleCol_v]])
dataClass_v <- class(subCell_dt[[sampleCol_v]])

if (metaClass_v != dataClass_v) {
  cat(sprintf("uniqCol: %s has mismatched classes.\n%s in meta\n%s in data.\nChanging both to character.\n",
              sampleCol_v, metaClass_v, dataClass_v))
  subMeta_dt[[sampleCol_v]] <- as.character(subMeta_dt[[sampleCol_v]])
  subCell_dt[[sampleCol_v]] <- as.character(subCell_dt[[sampleCol_v]])
}

### Determine if multiple entries
if (nrow(subMeta_dt) > 1) {
  multi_v <- T
  cat(sprintf("There are %s entries for this patient.\n", nrow(subMeta_dt)))
  myKable(subMeta_dt)
} else {
  multi_v <- F
  cat("There is 1 entry for this patient.\n")
  myKable(subMeta_dt)
}

### Determine if faceting (can this be done better)
facet_v <-  ifelse(length(which(is.na(subMeta_dt$facet))) == 0, T, F)

### Get plots from config table
plotCols_lsv <- list("stackedBar" = grep("[Ss]tacked[Bb]ar", colnames(config_dt), value = T),
                     "pie" = grep("[Pp]ie[Cc]hart", colnames(config_dt), value = T),
                     "sunburst" = grep("[Ss]unburst", colnames(config_dt), value = T),
                     "bar" = grep("[Bb]ar[Pp]lot", colnames(config_dt), value = T))

### Remove empties
plotLengths_v <- sapply(plotCols_lsv, length)
toPlot_v <- names(plotLengths_v[plotLengths_v != 0])
plotCols_lsv <- plotCols_lsv[toPlot_v]

```

```{r makeStacked, echo = F, fig.height= 7, fig.width=15}

if ("stackedBar" %in% names(plotCols_lsv)) {

  for (i in 1:length(plotCols_lsv$stacked)) {
    
    ### Get current column and subset config
    currPlotCol_v <- plotCols_lsv$stacked[i]
    currPlotName_v <- simpleCap(gsub("\\.", " ", gsub("^.*_", "", currPlotCol_v)))
    currConfig_dt <- config_dt[!is.na(get(currPlotCol_v)),mget(c("Class", currPlotCol_v))]
    currLevels_v <- currConfig_dt$Class[order(currConfig_dt[[currPlotCol_v]])]
    
    ### Subset data columns and merge metadata
    if (facet_v) {
      mergeCols_v <- c(sampleCol_v, "facet")
    } else {
      mergeCols_v <- sampleCol_v
    }
    
    currData_dt <- merge(subMeta_dt[,mget(mergeCols_v)],
                         subCell_dt[,mget(c(sampleCol_v, currConfig_dt$Class))], 
                         by = sampleCol_v, sort = F)
    
    #if (facet_v) setkey(currData_dt, "facet")
    
    ### Melt data
    currMelt_dt <- melt(currData_dt, id.vars = mergeCols_v)
    
    ### Subset colors
    currColor_dt <- color_dt[Population %in% currConfig_dt$Class,]
    
    ### Order output
    currMelt_dt$variable <- factor(currMelt_dt$variable, levels = currLevels_v)
    currColor_dt <- currColor_dt[match(currLevels_v, currColor_dt$Population)]
    
    ### Assign x variable
    if (facet_v) { x_v <- "facet" } else { x_v <- sampleCol_v}
    
    ### Get length of x-axis
    xLen_v <- length(unique(currMelt_dt[[x_v]]))
    
    ### Make plot
    curr_gg <- ggplot(data = currMelt_dt, aes(x = !!sym(x_v), y = value, fill = variable)) +
      geom_bar(position = "stack", stat = "identity") + 
      #my_theme() +
      big_label() +
      scale_fill_manual(values = currColor_dt$Hex, breaks = currColor_dt$Population) +
      ggtitle(currPlotName_v) + labs(y = bquote('Cell Density (cells / mm'^2*')'),
                                     fill = "Cell Type")
    
    ### Make table
    display_dt <- myT(currData_dt[,mget(c(x_v, currLevels_v))], newName_v = "Class")
    display_dt <- convertDFT(round(convertDFT(display_dt, col_v = "Class"), digits = 2), newName_v = "Class")
    display_grob <- tableGrob(display_dt, rows = NULL, theme = ttheme_default(base_size = 16))
    
    ### Output
    if (length(xLen_v > 5)) {
      foo <- grid.arrange(grobs = list(display_grob, curr_gg), nrow = 2)
    } else {
      foo <- grid.arrange(grobs = list(display_grob, curr_gg), ncol = 2)
    }
    
  }
  
}



```

```{r makePie, echo = F, fig.width=10}

if ("pie" %in% names(plotCols_lsv)) {

  for (i in 1:length(plotCols_lsv$pie)) {
  
  ### Get current column and subset config
  currPlotCol_v <- plotCols_lsv$pie[i]
  currPlotName_v <- simpleCap(gsub("\\.", " ", gsub("^.*_", "", currPlotCol_v)))
  currConfig_dt <- config_dt[!is.na(get(currPlotCol_v)),mget(c("Class", currPlotCol_v))]
  currLevels_v <- currConfig_dt$Class[order(currConfig_dt[[currPlotCol_v]])]
  
  ### Calculate percentages (and subset columns)
  currData_df <- convertDFT(subCell_dt[,mget(c(sampleCol_v, currConfig_dt$Class))], col_v = sampleCol_v)
  currData_df <- currData_df / rowSums(currData_df) * 100
  
  ### Merge metadata
  if (facet_v) {
    mergeCols_v <- c(sampleCol_v, "facet")
  } else {
    mergeCols_v <- sampleCol_v
  }
  
  currData_dt <- merge(subMeta_dt[,mget(mergeCols_v)],
                       convertDFT(currData_df, newName_v = sampleCol_v),
                       by = sampleCol_v, sort = F)
  
  if (facet_v) setkey(currData_dt, "facet")
  
  ### Melt
  currMelt_dt <- data.table::melt(currData_dt, id.vars = mergeCols_v)
  
  ### Subset colors
  currColor_dt <- color_dt[Population %in% currConfig_dt$Class,]
  
  ### Order output
  currMelt_dt$variable <- factor(currMelt_dt$variable, levels = currLevels_v)
  
  ### Make plot
  ### to do - non-facet option
  curr_lsgg <- lapply(unique(currMelt_dt$facet), function(x) {
    plot_gg <- ggplot(data = currMelt_dt[facet == x,], aes(x = facet, y = value, fill = variable)) +
      geom_bar(width = 1, stat = "identity") + pie_theme() +
      coord_polar(theta = "y", start = 0) +
      scale_fill_manual(values = currColor_dt$Hex, breaks = currColor_dt$Population) +
      ggtitle(x)
    return(plot_gg)
  })
  
  ### Get columns
  ncol_v <- ifelse(length(curr_lsgg) == 1, 1, 2)
  nrow_v <- ceiling(length(curr_lsgg) / 2)
  
  ### Grid plot
  curr_gg <- ggarrange(plotlist = curr_lsgg, ncol = ncol_v, nrow = nrow_v, common.legend = T, legend = "right")
  out_gg <- annotate_figure(p = curr_gg, top = text_grob(currPlotName_v, size = 24))
  
  ### Print
  print(out_gg)
  
}
  
}

```

\pretolerance=10000

```{r makeBar, echo = F, fig.width=20, fig.height=10}
# my_plot_hook <- function(x, options)
#   paste("\n", knitr::hook_plot_tex(x, options), "\n")
# knitr::knit_hooks$set(plot = my_plot_hook)

if ("bar" %in% names(plotCols_lsv)) {

  for (i in 1:length(plotCols_lsv$bar)) {
  
  ### Get current column and subset config
  currPlotCol_v <- plotCols_lsv$bar[i]
  currPlotName_v <- simpleCap(gsub("\\.", " ", gsub("^.*_", "", currPlotCol_v)))
  currConfig_dt <- config_dt[!is.na(get(currPlotCol_v)),mget(c("Class", "Gate", currPlotCol_v))]
  currConfig_dt <- currConfig_dt[get(currPlotCol_v) != "",]
  
  ### Grab which functional markers to plot (or all if not specified)
  currFuncMarkers_v <- gsub("Areap_|p$", "", currConfig_dt[get(currPlotCol_v) == "T",Gate])
  if (length(currFuncMarkers_v) == 0) currFuncMarkers_v <- setdiff(colnames(subFunc_lsdt[[1]]), "Class")
  
  ### Remove from config and make sure plot column is numeric
  currConfig_dt <- currConfig_dt[get(currPlotCol_v) != "T",]
  currConfig_dt[[currPlotCol_v]] <- as.numeric(currConfig_dt[[currPlotCol_v]])
  
  currLevels_v <- currConfig_dt$Class[order(currConfig_dt[[currPlotCol_v]])]
  
  ### Subset data
  currFunc_lsdt <- sapply(names(subFunc_lsdt), function(x) {
    
    y <- subFunc_lsdt[[x]]
    yy <- y[Class %in% currLevels_v,mget(c("Class", grep(paste(currFuncMarkers_v, collapse = "|"), colnames(y), value = T)))]
    if (facet_v) yy$facet <- subMeta_dt[get(sampleCol_v) == x, "facet"]
    return(yy)
    
  }, simplify = F, USE.NAMES = T)
  
  ### Combine
  currFunc_dt <- do.call(rbind, currFunc_lsdt)
  
  if (facet_v) setkey(currFunc_dt, "facet")
  
  ### Melt
  if (facet_v) {
    meltCols_v <- c("Class", "facet")
  } else {
    #meltCols_v <- c("Class", sampleCol_v)
    meltCols_v <- "Class"
  }
  
  currMelt_dt <- melt(currFunc_dt, id.vars = meltCols_v)

  
  ### Colors
  currMelt_dt$variable <- gsub("_func", "", currMelt_dt$variable)
  currColor_dt <- color_dt; currColor_dt[1,1] <- currColor_dt[1,1]
  currColor_dt$Population <- gsub("\\+", "", currColor_dt$Population)
  currColor_dt <- currColor_dt[Population %in% unique(currMelt_dt$variable),]
  
  ### Make plot
  curr_gg <- ggplot(data = currMelt_dt, aes(x = Class, y = value, fill = variable)) +
    geom_bar(stat = "identity", position = "stack") + my_theme() + angle_x() +
    scale_fill_manual(values = currColor_dt$Hex, breaks = currColor_dt$Population) +
    ggtitle(currPlotName_v) + labs(y = bquote('Marker Density (cells / mm'^2*')'),
                                   fill = "Functional Marker",
                                   x = "Cell Type")
  
  ### Add Facet
  if (facet_v) {
    curr_gg <- curr_gg + facet_wrap(~facet)
  }
  
  ### Make table
  round_dt <- round(currFunc_dt[,mget(setdiff(colnames(currFunc_dt), meltCols_v))], digits = 2)
  round_dt <- cbind(currFunc_dt[,mget(meltCols_v)], round_dt)
  display_grob <- tableGrob(round_dt, rows = NULL, theme = ttheme_default())
  
  ### Output
  #grid.arrange(grobs = list(display_grob, curr_gg), nrow = 2)
  
  grid.arrange(grobs = list(display_grob, curr_gg), ncol = 2)
  
}
  
}
```

# Sunburst

```{r makeSunburst, results='asis', echo = F}
#```{r, makeSunburst, echo = F, fig.height= 7, eval = T}
#plotCols_lsv$sunburst <- plotCols_lsv$sunburst[1:2]

full_lsls <- htmltools::tagList()
table_lsdt <- list()

if ("sunburst" %in% names(plotCols_lsv)) {

  for (i in 1:length(plotCols_lsv$sunburst)) {
  
  ### Get current column and name
  currPlotCol_v <- plotCols_lsv$sunburst[i]
  currPlotName_v <- simpleCap(gsub("\\.", " ", gsub("^.*_", "", currPlotCol_v)))
  
  ### Subset config and wrangle to get levels and match
  currConfig_dt <- config_dt[!is.na(get(currPlotCol_v)),mget(c("Class", currPlotCol_v))]
  currConfig_dt <- currConfig_dt[get(currPlotCol_v) != "",]
  currConfig_dt$Level <- as.numeric(gsub("[A-Z]*", "", currConfig_dt[[currPlotCol_v]]))
  currConfig_dt$Match <- gsub("[0-9]*", "", currConfig_dt[[currPlotCol_v]])
  currLevels_v <- unique(currConfig_dt$Level)
  
  ### Subset data columns and merge metadata
  currData_dt <- merge(subMeta_dt[,mget(c(sampleCol_v, "facet"))],
                       subCell_dt[,mget(c(sampleCol_v, currConfig_dt$Class))], 
                       by.x = sampleCol_v, by.y = sampleCol_v, sort = F)
  setkey(currData_dt, "facet") # sort by timepoint
  
  ### Transpose
  currForSData_dt <- myT(currData_dt[,mget(c("facet", currConfig_dt$Class))], newName_v = "Class")
  
  ### Have to wrangle for each timepoint individually
  currSData_lsdt <- list()
  
  for (j in 2:ncol(currForSData_dt)) {
    
    ### Subset class and current timepoint
    currName_v <- colnames(currForSData_dt)[j]
    curr_dt <- currForSData_dt[,c(1,j),with=F]
    
    ### Build base
    currSData_dt <- curr_dt[Class %in% currConfig_dt[Level == 1,Class]]
    currSData_dt$Parent <- NA
    currSData_dt$ID <- currSData_dt$Class
    
    ### For each subsequent level, grab parents, labels, etc. and then bind to initial table
    for (k in 2:length(currLevels_v)) {
      lvl_v <- currLevels_v[k]
      currNext_dt <- curr_dt[Class %in% currConfig_dt[Level == lvl_v,Class],]
      currNext_dt$Parent <- sapply(currNext_dt$Class, function(x) {
        m_v <- currConfig_dt[Class == x,Match]
        p_v <- currConfig_dt[Match == m_v & Level <= (lvl_v-1),Class]
        if (length(p_v) > 1) p_v <- paste(p_v, collapse = " - ")
        return(p_v)
      })
      currNext_dt$ID <- paste(currNext_dt$Parent, currNext_dt$Class, sep = " - ")
      currSData_dt <- rbind(currSData_dt, currNext_dt)
    } # for k
    
    ### Add column to map colors - need a better way to do this
    temp <- sapply(currSData_dt$Class, function(x) grep("KI67\\+|GRZB\\+|PD1\\+|CD163\\+|PDL1\\+", strsplit(x, split = " ")[[1]], 
                                                        value = T)[1], simplify = F)
    currSData_dt$colorMap <- sapply(temp, function(x) {if (length(x) == 0 | is.na(x)) { "" } else { x }})
    
    ### CHECK THIS! 
    ### Originally had functional markers as [marker]+, but now have them as just [marker].
    ### Some populations are called [marker]+ Cell Pop. and we remove the end to get [marker]+.
    ### We need to remove the + at the end so it matches. Is this the best way to do this, should
    ### we just put the + back in the color file?
    currSData_dt$colorMap <- gsub("\\+$", "", currSData_dt$colorMap)
    
    ### Change others to their class
    currSData_dt[colorMap == "", colorMap := Class]
    
    ### Add colors
    currSData_dt <- merge(currSData_dt, color_dt[,mget(c("Population", "Hex"))],
                          by.x = "colorMap", by.y = "Population", sort = F)
    
    ### Escape special char
    #currSData_dt$ID <- gsub("PD1\\-", "PD1\\\\\\-", currSData_dt$ID)
    
    ### Add to overal lsit
    currSData_lsdt[[currName_v]] <- currSData_dt
    
  } # for j
  
  ### Make plots
  currPlots_lspt <- sapply(names(currSData_lsdt), function(x) {
    plot_ly(data = currSData_lsdt[[x]], ids = ~ID, labels = ~Class, parents = ~Parent, marker = list(colors = ~Hex),
            values = as.formula(paste0("~", x)), type = "sunburst", branchvalues = "total") %>%
      layout(title = x)
  }, simplify = F, USE.NAMES = T)
  
  ### Make display tables
  temp_lsdt <- currSData_lsdt
  temp_lsdt <- sapply(temp_lsdt, function(x) {
    parents_v <- x[is.na(Parent),Class]
    newOrder_v <- unlist(sapply(parents_v, function(y) {
      grep(y, x$Class, fixed = T, value = T)
    }, simplify = F), use.names = F)
    out <- x[match(newOrder_v, x$Class)]
    out$colorMap <- NULL
    out$Parent <- NULL
    out$ID <- NULL
    out$Hex <- NULL
    return(out)
  }, simplify = F)
  
  if (length(temp_lsdt) == 1) {
    temp_dt <- temp_lsdt[[1]]
  } else {
    temp_dt <- tempMergeDTs(temp_lsdt, mergeCol_v = "Class")
  }

  ### Round
  cols_v <- setdiff(colnames(temp_dt), "Class")
  temp_dt[, (cols_v) := lapply(.SD, function(x) round(x, digits = 2)), .SDcols = cols_v]

  table_lsdt[[currPlotName_v]] <- temp_dt
  
  ### Wrap in tag list
  currPlots_lspt <- htmltools::tagList(currPlots_lspt)
  
  ### Add to other tag list
  full_lsls[[currPlotName_v]] <- currPlots_lspt
  
  }
  
  full_lsls
  
}



```

```{r, echo = F, results='asis', eval = T}

if ("sunburst" %in% names(plotCols_lsv)) {

  for (i in 1:length(table_lsdt)) {
    grid::grid.newpage()
    grid::grid.draw(tableGrob(table_lsdt[[i]], rows = NULL, theme = ttheme_default()))
    cat("\n\n\n")
  }

}
 
```